### 2025-06-10 simulation.qmd
20250610 - updates

```{r}
vars_by_pos <- list(
  QB = c(
    "games",
    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "fumbles_lost"#,
    #"sacks",
    #"pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",  
    #"pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_40_yds", "pass_400_yds",
    #"rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds"
    ),
  RB = c(
    "games",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds",
    "fumbles_lost",
    "return_yds", "return_tds"#,
    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
  ),
  WR = c(
    "games",
    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds",
    "fumbles_lost",
    "return_yds", "return_tds"#,
    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
  ),
  TE = c(
    "games",
    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds",
    "fumbles_lost",
    "return_yds", "return_tds"#,
    #"rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    #"rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
    #"rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds",
  )
)
```

```{r}
bootstrapSimulation <- function(projectedStats, vars_by_pos, n_iter = 10000, seed = NULL) {
  dt <- data.table::as.data.table(projectedStats)
  all_ids <- unique(dt$id)

@@

  future::plan(future::multisession)

  results <- future.apply::future_lapply(
    all_ids,
    function(player_id) {
      player_data <- dt[id == player_id]
      player_pos <- unique(player_data$pos)

@@

      stat_vars <- vars_by_pos[[player_pos]]
      out <- data.table(iteration = seq_len(n_iter), id = player_id, pos = player_pos)

      for (var in stat_vars) {
        if (var %in% names(player_data)) {
          non_na_values <- player_data[[var]][!is.na(player_data[[var]])]

          if (length(non_na_values) > 0) {
            out[[var]] <- sample(non_na_values, n_iter, replace = TRUE)
          } else {
            out[[var]] <- NA_real_
          }
        } else {
          out[[var]] <- NA_real_
        }
      }

      return(out)
    },
    future.seed = TRUE
  )

  data.table::rbindlist(results, use.names = TRUE, fill = TRUE)
}

bootstappedStats <- bootstrapSimulation(
  projectedStats = all_proj,
  vars_by_pos = vars_by_pos,
  n_iter = 10000,
  seed = 52242)

bootstappedStats$data_src <- bootstappedStats$iteration

@@

bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","team")],
    by = c("id" = "mfl_id")
  )
```

@@

```{r}
bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints %>% 
  group_by(id) %>% 
  summarise(
    mean = mean(raw_points, na.rm = TRUE),
    SD = sd(raw_points, na.rm = TRUE),
    min = min(raw_points, na.rm = TRUE),
    max = max(raw_points, na.rm = TRUE),
    range = max(raw_points, na.rm = TRUE) - min(raw_points, na.rm = TRUE),
    IQR = IQR(raw_points, na.rm = TRUE),
    MAD = mad(raw_points, na.rm = TRUE),
    CV = SD/mean,
    median = median(raw_points, na.rm = TRUE),
    pseudomedian = DescTools::HodgesLehmann(raw_points, na.rm = TRUE),
    mode = petersenlab::Mode(raw_points, multipleModes = "mean"),
    skewness = psych::skew(raw_points, na.rm = TRUE),
    kurtosis = psych::kurtosi(raw_points, na.rm = TRUE)
  )

bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints_summary %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","team","position")],
    by = c("id" = "mfl_id")
  ) %>% 
  select(name, team, position, mean:kurtosis, everything())

bootstrappedFantasyPoints_summary
```

```{r}
#| eval: false

>>>>

```{r}
vars_by_pos <- list(
  QB = c(
    "games",
    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "fumbles_lost",
    "sacks",
    "pass_09_tds", "pass_1019_tds", "pass_2029_tds", "pass_3039_tds", "pass_4049_tds", "pass_50_tds",  
    "pass_250_yds", "pass_300_yds", "pass_350_yds", "pass_40_yds", "pass_400_yds",
    "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds"
    ),
  RB = c(
    "games",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds",
    "fumbles_lost",
    "return_yds", "return_tds",
    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
  ),
  WR = c(
    "games",
    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds",
    "fumbles_lost",
    "return_yds", "return_tds",
    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
  ),
  TE = c(
    "games",
    "pass_att", "pass_comp", "pass_yds", "pass_tds", "pass_int",
    "rush_att", "rush_yds", "rush_tds",
    "rec_tgt", "rec", "rec_yds", "rec_tds",
    "fumbles_lost",
    "return_yds", "return_tds",
    "rush_09_tds", "rush_1019_tds", "rush_2029_tds", "rush_3039_tds", "rush_4049_tds", "rush_50_tds",
    "rush_50_yds", "rush_100_yds", "rush_150_yds", "rush_200_yds", "rush_40_yds",
    "rec_rz_tgt", "rec_50_yds", "rec_100_yds", "rec_150_yds", "rec_200_yds", "rec_40_yds"
  ),
  K = c(
    "fg_50", "fg_50_att", "fg_49", "fg_49_att", "fg_39", "fg_att_39", "fg", "fg_att", "xp", "xp_att"
  ),
  DST = c(
    "dst_fum_recvr", "dst_sacks", "dst_fumbles", "dst_tackles", "dst_td", "dst_yds_against", "dst_pts_against"
  ) ### ADD IDP VARS
)
```

```{r}
bootstrapSimulation <- function(projectedStats, vars_by_pos, n_iter = 10000, seed = NULL) {
  dt <- data.table::as.data.table(projectedStats) # using data.table for speed
  all_ids <- unique(dt$id)

@@

  future::plan(future::multisession) # parallelize tasks across multiple background R sessions using multiple cores to speed up simulation

  results <- future.apply::future_lapply(
    all_ids, # apply the function below to each player using a parallelized loop
    function(player_id) {
      player_data <- dt[id == player_id]
      player_pos <- unique(player_data$pos)

@@

      stat_vars <- vars_by_pos[[player_pos]] # pulls the relevant stat variables to simulate for this player's position
      out <- data.table(iteration = seq_len(n_iter), id = player_id, pos = player_pos)

      for (var in stat_vars) { # Loop over each stat variable that should be simulated for the player's position
        if (var %in% names(player_data)) {
          non_na_values <- player_data[[var]][!is.na(player_data[[var]])] # pulls non-missing values of the stat for the player (from all projection sources)

          if (length(non_na_values) > 0) {
            out[[var]] <- sample(non_na_values, n_iter, replace = TRUE) # If there are valid values, sample with replacement to simulate n_iter values
          } else {
            out[[var]] <- NA_real_ # specifies a numeric missing value (if all values were missing)
          }
        } else {
          out[[var]] <- NA_real_ # specifies a numeric missing value (if the stat variable doesn't exist)
        }
      }

      return(out)
    },
    future.seed = TRUE # ensures that each parallel process gets a reproducible random seed
  )

  data.table::rbindlist(results, use.names = TRUE, fill = TRUE) # combines all the individual player results into one large data table, aligning columns by name; fill = TRUE ensures that missing columns are filled with NA where necessary
}
```

```{r}
bootstappedStats <- bootstrapSimulation(
  projectedStats = all_proj,
  vars_by_pos = vars_by_pos,
  n_iter = 10000,
  seed = 52242)
```

```{r}
bootstappedStats$data_src <- bootstappedStats$iteration

@@

bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","team")],
    by = c("id" = "mfl_id")
  )

bootstrappedFantasyPoints <- bootstrappedFantasyPoints %>% 
  rename(projectedPoints = raw_points) %>% 
  arrange(-projectedPoints)
```

@@

```{r}
bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints %>% 
  group_by(id) %>% 
  summarise(
    mean = mean(projectedPoints, na.rm = TRUE),
    SD = sd(projectedPoints, na.rm = TRUE),
    min = min(projectedPoints, na.rm = TRUE),
    max = max(projectedPoints, na.rm = TRUE),
    range = max(projectedPoints, na.rm = TRUE) - min(projectedPoints, na.rm = TRUE),
    IQR = IQR(projectedPoints, na.rm = TRUE),
    MAD = mad(projectedPoints, na.rm = TRUE),
    CV = SD/mean,
    median = median(projectedPoints, na.rm = TRUE),
    pseudomedian = DescTools::HodgesLehmann(projectedPoints, na.rm = TRUE),
    mode = petersenlab::Mode(projectedPoints, multipleModes = "mean"),
    skewness = psych::skew(projectedPoints, na.rm = TRUE),
    kurtosis = psych::kurtosi(projectedPoints, na.rm = TRUE)
  )

bootstrappedFantasyPoints_summary <- bootstrappedFantasyPoints_summary %>% 
  left_join(
    nfl_playerIDs[,c("mfl_id","name","merge_name","team","position")],
    by = c("id" = "mfl_id")
  ) %>% 
  select(name, team, position, mean:kurtosis, everything()) %>% 
  arrange(-mean)
```

```{r}
bootstrappedFantasyPoints_summary %>% 
  filter(position == "QB") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))

bootstrappedFantasyPoints_summary %>% 
  filter(position == "RB") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))

bootstrappedFantasyPoints_summary %>% 
  filter(position == "WR") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))

bootstrappedFantasyPoints_summary %>% 
  filter(position == "TE") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))

bootstrappedFantasyPoints_summary %>% 
  filter(position == "PK") %>% 
  mutate(
    across(
      where(is.numeric),
      \(x) round(x, digits = 2)))
```
